---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# An R client for the BAN API

The `banR` package is a light R client for the [BAN API](https://adresse.data.gouv.fr/api/). The [Base Adresse Nationale (BAN)](https://adresse.data.gouv.fr/) is an open database of French adresses, produced by OpenStreetMap, La Poste, the IGN and Etalab. 

`banR` is not yet on CRAN, so for the time being it can be installed through `devtools`:

```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("joelgombin/banR")
```


`banR` allows to geocode lots of adresses in batch (the only hard limit is that, at the moment, the API only allows CSV files up to 8 MB). Please be gentle with the server though!

`banR` is designed to be used in a data exploration workflow, with a syntax 'à la `dplyr`':

```{r}
library(dplyr)
library(banR)
data("paris2012")

test_df <- paris2012 %>%
              slice(1:10) %>%
              mutate(adresse = paste(numero, voie, nom),
                     code_insee = paste0("751", arrondissement))
ban_search(test_df, adresse, code_insee = "code_insee") %>% glimpse

```

Please report issues and suggestions to the [issues tracker](https://github.com/joelgombin/banR/issues).
