---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# An R client for the BAN API

<!-- badges: start -->
[![R build status](https://github.com/joelgombin/banR/workflows/R-CMD-check/badge.svg)](https://github.com/joelgombin/banR/actions)
<!-- badges: end -->
  
The `banR` package is a light R client for the [BAN API](https://geo.api.gouv.fr/adresse). The [Base Adresse Nationale (BAN)](https://adresse.data.gouv.fr/) is an open database of French adresses, produced by OpenStreetMap, La Poste, the IGN and Etalab. 

`banR` can be installed from CRAN (stable version):

```{r eval=FALSE}
install.packages("banR")
```

or from Github (dev version):

```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("joelgombin/banR", build_vignettes = TRUE)
```


`banR` allows to geocode lots of adresses in batch (the only hard limit is that, at the moment, the API only allows CSV files up to 8 MB). Please be gentle with the server though!

`banR` is designed to be used in a data exploration workflow, with a syntax 'à la [`tidyverse`](http://tidyverse.org)':

```{r}
library(dplyr)
library(banR)
data("paris2012")

paris2012 %>%
  slice(1:100) %>%
  mutate(adresse = paste(numero, voie, nom),
         code_insee = paste0("751", arrondissement)) %>% 
  geocode_tbl(adresse = adresse, code_insee = code_insee) %>%
  glimpse

```
To know more about this package, please read the [vignette](http://joelgombin.github.io/banR/articles/geocode.html) (`vignette("geocode")`)

Please report issues and suggestions to the [issues tracker](https://github.com/joelgombin/banR/issues).

## See also

* [BAN-geocoder](https://github.com/atao/BAN-Geocoder), python wrapper for adresse.data.gouv.fr
