---
title: "Geocoding French adresses with BanR"
author: "Paul-Antoine Chevalier (Etalab), Joël Gombin (Datactivist)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Geocoding French adresses with BanR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
library("tibble")
library("dplyr")
library("banR")
# generate fake data
table_test <- tibble::tibble(
  adress = c("39 quai André Citroën", "64 Allée de Bercy", "20 avenue de Ségur"),
  postal_code = c("75015", "75012", "75007"),
  z = rnorm(3)
  )
```

- `geocode()` geocodes a single address
- `reverse_geocode()` reverse geocodes a single pair of longitude and latitude
- `geocode_tbl()` geocodes a data frame
- `reverse_geocode_tbl()` reverse geocodes a data frame

## Geocode

Geocoding is the process of transforming a human readable address into a location (ie a pair of latitude and longitude).

### A single address

```{r geocode}
geocode(query = "39 quai André Citroën, Paris") %>%
  glimpse()
```

The BAN API sends back both projected/Cartesian coordinates (`x` and `y` columns - they use Lambert 93 projection, aka as EPSG:2154), and lon/lat (i.e. WGS84) coordinates (`longitude` and `latitude` columns). It also indicates the degree of confidence it has in each result (column `score`). The above example only sends back one result, but sometimes the API will send back several suggestion for the same query. They are ordered by descending order of confidence. 

### A data frame

In addition to the adress, `geocode_tbl()` can take as argument either the [postal code](https://en.wikipedia.org/wiki/Postal_codes_in_France) or the French official code ([INSEE code](https://en.wikipedia.org/wiki/INSEE_code)) of the commune. 

```{r geocode-tbl}
geocode_tbl(tbl = table_test, adresse = adress) %>%
  glimpse()
```

```{r geocode-tbl-postalcode}
geocode_tbl(tbl = table_test, adresse = adress, code_postal = postal_code) %>%
  glimpse()
```

```{r geocode-tbl-codeinsee}
data("paris2012")
paris2012 %>%
  slice(1:100) %>%
  mutate(
    adresse = paste(numero, voie, nom),
    code_insee = paste0("751", arrondissement)
    ) %>%
  geocode_tbl(adresse = adresse, code_insee = code_insee) %>%
  glimpse()
```

## Reverse geocode

Reverse geocoding is the process of back (reverse) coding of a point location (latitude, longitude) to a human readable address.

### A single adress

`reverse_geocode()` takes longitude and latitude as arguments and returns a data frame with addresses.

```{r reverse-geocode}
reverse_geocode(long =  2.279092, lat = 48.84683)  %>%
  glimpse()
```

### A data frame

`reverse_geocode_tbl` takes the names of the longitude and latitude columns and returns a data frame with adresses.

```{r reverse-geocode-tbl}
test_df <- tibble::tibble(
  nom = sample(letters, size = 10, replace = FALSE),
  lon = runif(10, 2.19, 2.47),
  lat = runif(10, 48.8, 48.9)
)

test_df %>% 
  reverse_geocode_tbl(lon, lat) %>% 
  glimpse

```


### Geocode from a CSV file: 


`geocode_csv()` allows you, from a csv file, to geocode a massive amount of addresses. __The API can handle a csv file up to 50Mo__. In order to use the function, you'll need a local CSV file. 

In this example we'll work with the CSV file provided by the API DOC website, available here: __https://adresse.data.gouv.fr/exemples/search.csv__

```{r}

# 1.first download the csv file 

download.file("https://adresse.data.gouv.fr/exemples/search.csv", 
              "search.csv")

# 2. geocode the addresses 

geocode_csv(file = "search.csv")


```

__By default `geocode_csv()` will concatenate all the columns in order to formulate the addresses to be geocoded and it will return a tibble containing the geocoded addresses among other related information.__


You can also tune your geocoding by providing additional parameters to the `...` argument of the `geocode_csv()` function. For example, it's possible to select the columns that will be used for our geocoding: 


```{r}


geocode_csv(file = "search.csv", 
            columns = "postcode", 
            columns = "adresse")


```

In case you've a large CSV file and you want to speed up the process, you can indicate the needed resulting columns this way: 


```{r}

geocode_csv(file = "search.csv", 
            columns = "postcode", 
            columns = "adresse", 
            result_columns = "latitude", 
            result_columns = "longitude")


```

Finally, it's possible to indicate the column that contains the __INSEE code__ or the __Postal code__. The example CSV file that we're using doesn't contain these kind of columns but you'd implement it this way:

```{r, eval=FALSE}

# DO NOT RUN. THIS IS HYPOTHETICAL 

geocode_csv(file = "YOUR_CSV_FILE", 
            postcode = "COLUMN_WITH_POSTCODE", 
            citycode = "COLUMN_WITH_INSEECODE")

```

## 4. Reverse Geocode from a CSV file: `geocode_csv_reverse()`


`geocode_csv_reverse()` works the way as `geocode_csv()` except that it does the inverse. It transforms longitude and latitude information into addresses. __Note also that API tolerates CSV files up to 6Mo for this kind of operation.__ 

##### Important 🐱 The csv file must contain columns named _latitude (or lat)__ and _longitude (or lon or lng)_


In order to illustrate this function, we'll work on the CSV file provided by the API DOC and available here: __https://adresse.data.gouv.fr/exemples/reverse.csv__



```{r}

# 1.first download the csv file 

download.file("https://adresse.data.gouv.fr/exemples/reverse.csv", 
              "reverse.csv")

# 2. geocode the addresses 

geocode_csv_reverse(file = "reverse.csv")


```