---
title: "Geocoding French adresses with BanR"
author: "Paul-Antoine Chevalier (Etalab), Joël Gombin (Datactivist)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Geocoding French adresses with BanR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
library("tibble")
library("dplyr")
library("banR")
# generate fake data
table_test <- tibble::tibble(
  adress = c("39 quai André Citroën", "64 Allée de Bercy", "20 avenue de Ségur"),
  postal_code = c("75015", "75012", "75007"),
  z = rnorm(3)
  )
```

- `geocode()` geocodes a single address
- `reverse_geocode()` reverse geocodes a single pair of longitude and latitude
- `geocode_tbl()` geocodes a data frame
- `reverse_geocode_tbl()` reverse geocodes a data frame

## Geocode

Geocoding is the process of transforming a human readable address into a location (ie a pair of latitude and longitude).

### A single address

```{r geocode}
geocode(query = "39 quai André Citroën, Paris") %>%
  glimpse()
```

The BAN API sends back both projected/Cartesian coordinates (`x` and `y` columns - they use Lambert 93 projection, aka as EPSG:2154), and lon/lat (i.e. WGS84) coordinates (`longitude` and `latitude` columns). It also indicates the degree of confidence it has in each result (column `score`). The above example only sends back one result, but sometimes the API will send back several suggestion for the same query. They are ordered by descending order of confidence. 

### A data frame

In addition to the adress, `geocode_tbl()` can take as argument either the [postal code](https://en.wikipedia.org/wiki/Postal_codes_in_France) or the French official code ([INSEE code](https://en.wikipedia.org/wiki/INSEE_code)) of the commune. 

```{r geocode-tbl}
geocode_tbl(tbl = table_test, adresse = adress) %>%
  glimpse()
```

```{r geocode-tbl-postalcode}
geocode_tbl(tbl = table_test, adresse = adress, code_postal = postal_code) %>%
  glimpse()
```

```{r geocode-tbl-codeinsee}
data("paris2012")
paris2012 %>%
  slice(1:100) %>%
  mutate(
    adresse = paste(numero, voie, nom),
    code_insee = paste0("751", arrondissement)
    ) %>%
  geocode_tbl(adresse = adresse, code_insee = code_insee) %>%
  glimpse()
```

## Reverse geocode

Reverse geocoding is the process of back (reverse) coding of a point location (latitude, longitude) to a human readable address.

### A single adress

`reverse_geocode()` takes longitude and latitude as arguments and returns a data frame with addresses.

```{r reverse-geocode}
reverse_geocode(long =  2.279092, lat = 48.84683)  %>%
  glimpse()
```

### A data frame

`reverse_geocode_tbl` takes the names of the longitude and latitude columns and returns a data frame with adresses.

```{r reverse-geocode-tbl}
test_df <- tibble::tibble(
  nom = sample(letters, size = 10, replace = FALSE),
  lon = runif(10, 2.19, 2.47),
  lat = runif(10, 48.8, 48.9)
)

test_df %>% 
  reverse_geocode_tbl(lon, lat) %>% 
  glimpse

```